Numerical trajectory patterns of floating macroalgae

### 2021-01-11 thesis_chapter_3.Rmd
one error

Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape, however no research exists attempting to clarify the drifting characteristics of solitary macroalgae. Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes. The findings from this study indicate that solitary floating virtual *E. maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely lagrangian particle. Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies. Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are neglible when investigating dispersal patterns of drifting *E.maxima*. In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory but also causes higher density of particles to become entrained in an eddy. Overall the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not. This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately. Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these vary seasonally. The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.

>>>>

Most of the past research has focused on large macroalgae rafts which can vary greatly in size and shape, however no research exists attempting to clarify the drifting characteristics of solitary macroalgae. Understanding how to accurately model the distribution of solitary kelp can lead to be understanding of accumulation zones and sinks and ultimately the ecological links pertaining to those processes. The findings from this study indicate that solitary floating virtual *E. maxima* particles tend to follow the prevailing surface currents and the inclusion of wind causes the trajectories to differ greatly from that of a purely lagrangian particle. Furthermore, the inclusion of both hydrodynamic and wind drag cause clustering of particles along the trajectory as well causing particles to become entrained in eddies. Different cross-sectional areas exposed to hydrodynamic and wind drag had no effect on overall trajectory suggesting that those differences are negligible when investigating dispersal patterns of drifting *E.maxima*. In addition, the use of a cylindrical shape coefficient has no effect on mean trajectory but also causes higher density of particles to become entrained in an eddy. Overall the inclusion of drag forces is an important aspect to consider when investigating the dispersal patterns of solitary macroalgae, while differences in cross-sectional areas are not. This study also provides an approach which can be adapted to model any floating solitary macroalgae, provided the surface area can be estimated accurately. Gaps in the research exist when considering what oceanographic processes play a role in solitary macroalgal dispersal as well as how these vary seasonally. The identification of biological and physical factors that play a role in accumulation zones is also needed, as well as better estimates for sinking rates and raft-times.
